### 2014-03-10 src/docs/papers/agrad/stan-autodiff.tex
first version of start of agrad paper

\documentclass[10pt]{article}
\usepackage{stan-papers}

\title{\Large\bfseries \code{\bfseries agrad}: A C++ Library for Efficient, Scalable,
  and Extensible Automatic Differentiation}
\author{\large Bob Carpenter \\ {\small Columbia University}
        \\[8pt]
        \large Marcus Brubaker \\ {\small Toyota Technical Institute}
   \and \large Matt Hoffman \\ {\small Adobe Research Labs}
        \\[8pt]
        \large Peter Li \\ {\small Columbia University}
   \and \large Daniel Lee \\ {\small Columbia University}
        \\[8pt]
        \large Michael Betancourt \\ {\small University of Warwick}
}
\date{\vspace*{8pt}\normalsize \today}

\begin{document}
\maketitle

\begin{abstract} 
  \noindent
  The \code{agrad} C++, forward- and reverse-mode automatic
  differentiation library was defined to satisfy
  six competing goals: extensibility, efficiency, scalability,
  stability, portability, and usability. Its underlying
  object-oriented design
  and functional template abstraction supports the addition of new
  functions based on value and first-order derivative calculations
  abstracted from the underlying automatic differentiation mechanisms.
  It uses a custom memory arena and template metaprogramming for
  efficiency.  Its underlying reverse-mode 
  evaluation is lazy for both efficiency and scalability.
  First- and higher-order derivatives are avaialable through a
  functional abstraction which hides all of the details of the
  library's internals. For broad applicability and numerical
  stability, it implements an extensive library of special functions,
  matrix and linear algebra operations, probability functions, 
  and ordinary differential equation solvers.  Evaluations on test
  problems show that both its efficiency and scalability are state of
  the art.  The \code{agrad} library is open-source licensed under the
  new BSD license and has been tested for all major compilers for
  Windows, Mac OS X, and Linux. 
\end{abstract}

\section{Introduction}

Many scientific computing applications, including optimization,
simulation, and differential equation solvers, require the computation
of first-order and higher-order derivatives over functions defined on
an application-by-application basis.

Although derivatives can be calculated and coded by hand, the process
is both tedious and error prone, especially for algorithms represented
in large amounts of code. Two popular alternatives are finite
differencing and automatic differentiation. Finite differencing calculates a
derivative of a function $f$ at $x$ by the linear approximation
\[ 
\frac{d}{dx}f(x) 
\ \approx \
\frac{f(x + \epsilon) - f(x - \epsilon)}{2\epsilon}.
\]
Automatic differentiation directly differentiates a computer program
for $f$ by applying the chain rule to the expressions produced by the
program's statements.

In contrast to finite differencing methods, automatic differentiation
is efficient and accurate. As an example of efficiency, gradient
calculations require only a constant multiple of the time taken to
evaluate a function independent of dimension, whereas finite
differences require a multiple determined by the dimensionality.
Automatic differentiation computes derivatives up to machine
precision, whereas finite differences are notoriously imprecise
because they require a tradeoff of arithmetic precision and accuracy
in linear approximations that depends on scale and rarely allows more
than 1e-6 relative accuracy even with scale-based tuning.

The \code{agrad} library was introduced to overcome several
shortcomings of existing implementations such as CppAD, Saccado,
AdolC, and Adept. The main motivation was to allow easy extensibility
to enable the implementation of a wide range of special functions,
probability density, mass and cumulative distribution functions, and
matrix and linear-algebra operations. A secondary motivation was to
increase scalability and efficiency, particularly for the calculations
of log probability functions as required for statistical inference
through optimization and sampling. A third motivation was to provide
simple interfaces that were portable and required no
platform-dependent installation and could be both used and extended to
new functions without understanding the internal differentiation data
types. The rest of the paper describes how \code{agrad} is designed
and implemented and why it improves the state of the art in
efficiency, scalability, and usability.

\section{Derivative Calculation Interface}

The \code{agrad} library application programmer interface (API) is
designed to allow client code to calculate derivatives and extend the
built-in library without needing to know anything about the internals
of the library's underlying automatic differentiation data types.
This section describes how to calculate derivatives and the next how
to add user-defined functions.

To use the \code{agrad} library to calculate derivatives, client code
need only define a templated functor, define an argument vector,
define the appropriate vector or matrix to hold the result, and then
call the derivative functional of the apporpriate order.

The library also exposes an interface for direct access to the
variables and methods used for automatic differentiation for
finer-grained control and integration with other software.

\subsection{Example Function: Neal's Funnel}

The log of the funnel probability density function is a simple example
of the kinds of probability functions \code{agrad} was designed to
differentiation.  
kinds of functions
Consider the log of the funnel density
\citep{Neal:2003}, which is defined for
a positive scalar $v$ and $N$-vector $y$ by
\[
\mbox{funnel}(v,y) = \frac{1}{2} \left(
  N v
  + \frac{v^2}{9}
  + \exp(-v) \sum_{n=1}^N y_n^2
\right)
\]
\reffigure{funnel-cpp} provides a C++ implementation of the
funnel function suitable for use with \code{agrad}.
%
\begin{figure}
\begin{quote}
\begin{Verbatim}
struct funnel {
  template <typename T>
  T operator()(const std::vector<T>& x) const {
    T v = x[0];
    T sum_y_sq = 0;
    for (size_t n = 1; n < x.size(); ++n)
      sum_y_sq += x[n] * x[n];
    T p1 = N * v;
    T p2 = (v * v) / 9;
    T p3 = sum_y_sq * exp(-v);
    return (p1 + p2 + p3) / 2;
  }
};
\end{Verbatim}
\end{quote}
\vspace*{-6pt}
\caption{\it The funnel density implemented as a C++ functor operating
  on a C++ standard template library vector containing $v$ followed by
  $y$. The two quadratic terms could also be expressed using \code{agrad}'s
  built-in functions\, \code{square} or\, \code{pow}.  The function
  could alternatively be defined in terms of any class definining a
  \code{size()} method and allowing \code{operator[]} indexing, such 
  as an Eigen library vector or C++11 standard library array.
}\label{funnel-cpp.figure}
\end{figure}
%

\subsection{Derivative Calculation}

Derivatives can be calculated using either C++ functionals operating
on functions defined as functors, or directly using the underlying
automatic differentiation data types.

\subsubsection{Functional Approach}

The \code{agrad} library supports a functional approach to calculating
derivatives which is most easily explained with an example. The
following code calculates the value \code{fx} and gradient \code{grad}
of the funnel functor \code{f} evaluated at the argument vector
\code{x}.
%
\begin{quote}
\begin{Verbatim}
funnel f;
std::vector<double> x = ...;
double fx;
std::vector<double> grad;
agrad::gradient(f,x,fx,grad);
\end{Verbatim}
\end{quote}

\subsubsection{Direct Approach}

The \code{agrad} library also supplies the same kind of low-level
methods as other automatic differentiation libraries such as Adept,
CppAD, and Saccado, although without the need for turning on or off
``recording'' to explicitly manage memory. For example, the following
function defines the log of the Beta density function
%
\begin{eqnarray*}
f(x,\alpha,\beta) 
& = & \log \distro{Beta}(x|\alpha,\beta)
\\
& = & (\alpha - 1) \log x + (\beta - 1) \log (1 - x) 
  \log \Gamma(\alpha + \beta) - \log \Gamma(\alpha) - \log
  \Gamma(\beta),
\end{eqnarray*}
%
which can be implemented using built-in special functions as
%
\begin{quote}
\begin{Verbatim}
agrad::var f(double x, 
             const agrad::var& a, const agrad::var& b) {
  return (a - 1) * log(x) + (b - 1) * log1m(x)
         + lgamma(a + b) - lgamma(a) - lgamma(b);
}
\end{Verbatim}
\end{quote}
%
To calculate the derivatives of $y = f(x,\alpha,\beta)$ with respect to
$\alpha,\beta$ at
$\alpha = 1.9$, $\beta = 17.2$, and  $x = 0.329$, the code is
%
\begin{quote}
\begin{Verbatim}
agrad::var a = 1.9;    agrad::var b = 17.2;
double x = 0.329;
agrad::var y = f(x,a,b);
fab.grad();
double dy_da = a.adjoint();    double dy_db = b.adjoint();
\end{Verbatim}
\end{quote}
%
This approach can be used for arbitrary functions defined for
arguments including \code{var}; see \refsection{flexible-signatures}
for an example of how the beta density function is defined in
\code{agrad}.

\section{Flexible Type Signatures}\label{flexible-signatures.section}

To allow for maximum flexibility and efficiency in calling functions, 
\code{agrad}'s signature for the log Beta density is declared as
%
\begin{quote}
\begin{Verbatim}
template <typename Ty, typename Ta, typename Tb>
inline typename agrad::return_type<Ty,Ta,Tb>::type
beta_log(const Ty& y, const Ta& a, const Tb& b); 
\end{Verbatim}
\end{quote}
%
where a mechanism like that of Boost library's promotion template
metaprogram is used to calculate the return type automatically as the
most general type able to hold the result (e.g., \code{agrad::var} if
used in place of \code{f} in the previous example). In general,
\code{agrad}'s density functions are vectorized to allow any argument
to be a scalar, standard vector, or Eigen vector of any primitive type
or of type \code{var} (or of type \code{fvar<T>}, the forward-mode
automatic differentiation variable type).  The need for this
flexibility is why a simple log density library like Boost's, which
template on a single type, cannot be directly used for \code{agrad}.
%
\footnote{For maximum portability, the \code{agrad} library does not
  depend on C++11 extensions, but even if it did, the \code{auto}
  mechanism is not powerful to determine the result type here.}

\section{Reverse-Mode Automatic Differentiation}

\appendix

\section{Functional Derivative API}

The high-level \code{agrad} API for calculating derivatives is defined
in terms of C++ functionals, i.e., functions operating on functions.%
%
\footnote{To simplify notation, \code{Vector<T>} is used as shorthand
for \code{Eigen::Matrix<T,Eigen::Dynamic,1>} and 
\code{Matrix<T>} for \code{Eigen::Matrix<T,Eigen::Dynamic,Eigen::Dynamic,1>}.}

\subsection{Simple Derivatives}

If \code{f} is a functor implementing the signature
%
\begin{quote}
\begin{Verbatim}
template <typename T>
T operator()(const T& x) const;
\end{Verbatim}
\end{quote}
%
and \code{x} is a point,
%
\begin{quote}
\begin{Verbatim}
double x = ...;  
\end{Verbatim}
\end{quote}
%
then
%
\begin{quote}
\begin{Verbatim}
double fx, df_dx;  
agrad::derivative(f,x,fx,df_dx);
\end{Verbatim}
\end{quote}
%
sets \code{fx} to be the value of \code{f} at \code{x} and
sets \code{df\_dx} to be the derivative of \code{f} at \code{x}.

\subsection{Gradients and Hessians}

If \code{f} is a functor implementing the signature
%
\begin{quote}
\begin{Verbatim}
template <typename T>
T operator()(const Vector<T>& x) const;
\end{Verbatim}
\end{quote}
%
and \code{x} is a vector of inputs,
%
\begin{quote}
\begin{Verbatim}
Vector<double> x = ...;
\end{Verbatim}
\end{quote}
%
then
%
\begin{quote}
\begin{Verbatim}
double fx;    Vector<double> g;
agrad::gradient(f,x,fx,grad);
\end{Verbatim}
\end{quote}
%
sets \code{fx} to be the value of \code{f} at \code{x} and sets
\code{grad} to be the gradient of \code{f} at \code{x} and
%
\begin{quote}
\begin{Verbatim}
double fx;    Matrix<double> H;
agrad::hessian(funnel(),x,H);
\end{Verbatim}
\end{quote}
%
sets \code{fx} to be the value of \code{f} at \code{x} and sets
\code{H} to be the Hessian of \code{f} at \code{x}.

\subsection{Directional Gradients and Hessians}

There are also directional forms of gradients and Hessians, so that if
\code{v} is a vector representing a direction, 
%
\begin{quote}
\begin{Verbatim}
Vector<double> v = ...;
\end{Verbatim}
\end{quote}
%
then
%
\begin{quote}
\begin{Verbatim}
double fx;    double grad_v;
agrad::gradient_dot_vector(f,x,v,fx,grad_v);
\end{Verbatim}
\end{quote}
%
sets \code{fx} to be the value of \code{f} at \code{x} and sets
\code{grad\_v} to be the product of the gradient of \code{f} at
\code{x} and the vector \code{v}, and
%
\begin{quote}
\begin{Verbatim}
double fx;    Vector<double> H_v;
agrad::gradient_dot_vector(f,x,v,fx,Hv);
\end{Verbatim}
\end{quote}
%
sets \code{fx} to be the value of \code{f} at \code{x} and sets
\code{H\_v} to be the product of the Hessian of \code{f} at
\code{x} and the vector \code{v}.

\subsection{Jacobians}

If \code{f} is a functor implementing the signature
%
\begin{quote}
\begin{Verbatim}
template <typename T>
Vector<T> operator()(const Vector<T>& x) const;
\end{Verbatim}
\end{quote}
%
and \code{x} is an input vector, then
%
\begin{quote}
\begin{Verbatim}
VectorXd x = ...;
\end{Verbatim}
\end{quote}
%
then
%
\begin{quote}
\begin{Verbatim}
VectorXd fx;    MatrixXd J;
agrad::jacobian(f,x,fx,J);
\end{Verbatim}
\end{quote}
%
sets \code{fx} to be the value of \code{f} at \code{x} and 
\code{J} to be the Jacobian of \code{f} at \code{x}.

\bibliographystyle{apalike}
\bibliography{../../bibtex/all}

\end{document}