### 2014-12-01 src/docs/papers/stan-language/cut.tex

\section[Why Stan?]{Why \proglang{Stan}?}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

We did not set out to build \proglang{Stan} as it currently exists.%
%
\footnote{\proglang{Stan}'s home page, \url{http://mc-stan.org/},
  links to the \proglang{Stan} manual, example models including translations of
  the \proglang{BUGS} examples and the models from \citep{GelmanHill:2007},
  getting started instructions and full documentation for \proglang{Stan}'s
  interfaces for the command line shell (\pkg{CmdStan}), \proglang{Python}
  (\pkg{PyStan}), and \proglang{R} (\pkg{RStan}), and the source code repositories
  and issue trackers.}
%
Our original goal was to apply full Bayesian inference to the sort of multilevel
generalized linear models discussed in Part~II of
\citep{GelmanHill:2007}, which are structured with grouped and
interacted predictors at multiple levels, hierarchical covariance
priors, nonconjugate coefficient priors, latent effects as in
item-response models, and varying output link functions and
distributions.

These models turned out to be a challenge for existing general purpose
inference software.  A direct encoding in \proglang{BUGS}
\citep{LunnEtAl:2000,LunnEtAl:2009,LunnEtAl:2012} or
\proglang{JAGS}~\citep{Plummer:2003} can grind these tools to a halt.
We began by adding custom vectorized logistic regressions to
\proglang{JAGS} using \proglang{C++} to overcome the cost of
interpretation.  Although this is much faster than looping in
\proglang{JAGS}, it quickly became clear that the root of the problem
was the slow convergence of conditional sampling when parameters were
highly correlated in the posterior, as for time-series models and
hierarchical models with interacted predictors.  We finally realized
we needed a better sampler, not a more efficient implementation of
Gibbs sampling.

We briefly considered trying to tune proposals for a random-walk
Metropolis-Hastings sampler, but that seemed too problem specific and
not even necessarily possible without some kind of adaptation rather
than a global tuning of the proposals.

We were hearing more about Hamiltonian Monte Carlo (HMC) sampling,
which appeared promising but was also problematic in that the
Hamiltonian dynamics simulation requires the gradient of the log
posterior.  Although not difficult in theory for most functions,
computing these gradients by hand on a model-by-model basis is very
tedious and error prone.  That is when we discovered reverse-mode
algorithmic differentiation, which, given a templated \proglang{C++}
function for the log posterior, automatically computes its analytic
gradient up to machine precision accuracy in only a few multiples of
the time to evaluate the log probability function itself.%
%
\footnote{Hessian matrices (i.e., all second order derivatives) are
  more expensive to calculate; each row corresponds to a parameter and
  is filled by the gradient of the derivative of the log probability
  function with respect to the parameter.}
%
We explored existing algorithmic differentiation packages with open
licenses such as \pkg{RAD} \citep{Gay:2005} and its repackaging in the
Sacado module of the \pkg{Trilinos} toolkit, and the \pkg{CppAD}
package of the COIN-OR toolkit \citep{BellBurke:2008}.  Both packages,
however, supported few special functions (e.g., probability functions,
log gamma, inverse logit) or linear algebra operations (e.g., Cholesky
decompositions, matrix division), and neither are easily or modularly
extensible.

Consequently we built our own reverse-mode algorithmic differentiation package.
At that point, we ran into the problem that we could not just plug in
the probability functions from a package like \pkg{Boost} because they
weren't templated generally enough across all arguments.  Rather
than pay the price of promoting floating point values to algorithmic
differentiation variables, we wrote our own fully templated
probability functions and other special functions.

Next, we integrated the \proglang{C++} package \pkg{Eigen} for matrix
operations and linear algebra functions.  \pkg{Eigen} makes extensive
use of expression templates for lazy evaluation and the curiously
recurring template pattern (CRTP) to implement concepts without
virtual function calls; \cite{VandevoordeJosuttis:2002} provide a
complete description of template metaprogramming techniques.
Unfortunately, we ran into the same problem with Eigen as with the
existing probability libraries---it doesn't support mixed operations
of algorithmic differentiation variables and primitives like
\code{double}.  Although we initially began by promoting
floating-point vectors to algorithmic differentiation variables, we
later completely rewrote mixed-mode operations like multiplying a data
matrix by a parameter vector.  To compute derivatives of matrix
operations such as division and inverse, we implemented specialized
derivatives using the rules described by \cite{Giles:2008}.

At this point, we could fit models coded directly in \proglang{C++} on
top of the pre-release versions of the \proglang{Stan} application
programming interface (API).  Seeing how well this all worked, we set
our sights on the generality and ease of use of \proglang{BUGS} and
designed a modeling language in which statisticians could write their
models in familiar notation that could be transformed to efficient
\proglang{C++} code before being compiled into an efficient executable
program.  Although we started with the target of specifying directed
graphical models, our first implementation turned out to be a more
general imperative form of specifying log probability functions;
\proglang{BUGS} models could be translated line for line, but the
\proglang{Stan} language also supports much richer imperative
constructs like conditionals, while loops, and local variables.  This
paper is primarily about the \proglang{Stan} language and how it can
be used to specify statistical models.

The next problem we ran into when we started implementing richer
models was variables with constrained support, such as positive variables, 
simplexes, and covariance matrices.  
Efficient implementation of these constraints
required the introduction of typed variables which automatically transformed to
unconstrained support with suitable adjustments to the log probability
from the log absolute Jacobian determinant of the inverse transforms.

Even with the prototype compiler generating models, we still faced a
major hurdle for ease of use. The efficiency of HMC is very sensitive
to two tuning parameters, the discretization interval (i.e., step
size) and the total simulation time (i.e., number of steps).  The
interval size parameter could be tuned during warm-up based on
Metropolis rejection rates, but the number of steps proved difficult
to tune without sacrificing the detailed balance of the sampler.  This
led to the development of the No U-Turn (NUTS) sampler
\citep{HoffmanGelman:2011}.  Roughly speaking, NUTS builds a tree of
possible samples by randomly simulating Hamiltonian dynamics both
forwards and backwards in time until the combined trajectory turns
back on itself.  Once the trajectory has terminated, a new sample is
drawn from the tree.  In its original version, NUTS also estimates a
step size during warmup based on a target acceptance probability.

We thought we were home free, but when we measured the
speed of some \proglang{BUGS} examples versus \proglang{Stan} we were
very disappointed.  \proglang{BUGS}'s very first example model, Rats,
ran more than an order of magnitude faster in \proglang{JAGS} than in
\proglang{Stan}.  Rats is a tough test case because the conjugate
priors and lack of posterior correlations make it an ideal candidate
for efficient Gibbs sampling.  Realizing that we were doing redundant
calculations, we wrote a vectorized form of the normal distribution
for multiple variates with the same mean and scale, which sped things
up a bit.  Performance was lacking until we finally figured out how
to both eliminate redundant calculations and partially evaluate the
gradients using a combination of expression templates and
metaprogramming.

When we attempted to fit a time-series model, we found that
normalizing the data to unit sample mean and variance sped up the fits
by an order of magnitude.  In hindsight this isn't surprising, as the
performance of the numerical simulation in HMC scales with the
variation of the parameter scales.  In \proglang{Stan} 1.0 we
introduced an adaptive diagonal metric (mass matrix) into our HMC
implementations that allowed the parameter scales to be standardized
automatically; \proglang{Stan} 2.0 added an option to estimate a dense metric
(mass matrix) and a full covariance standardization of the posterior.

Because the diagonal and dense metrics perform only a global
standardization, the performance of \proglang{Stan} in models where
the relative parameter scales varies by location, such as hierarchical
and latent models, can suffer (hierarchical modeling is covered in
Section~\ref{hierarchical-modeling.section} and reparameterizations to speed it
up in Section~\ref{centering.section}).  Riemannian manifold Hamiltonian Monte
Carlo (RMHMC) introduces a location-dependent metric that can overcome
these final hurdles \citep{GirolamiCalderhead:2011}; \proglang{Stan} has a
prototype implementation of RMHMC based on the SoftAbs metric of
\citep{Betancourt:2012}, using a generalization of NUTS to Riemannian
manifolds \citep{Betancourt:2013}.

=============================================

\subsection{Variable definition and block execution summary}

A table summarizing the point at which variables are read, written,
or defined is provided in Figure~\ref{block-actions.fig}. 
%
\begin{figure}
\begin{center}
\begin{tabular}{l|c|l|l}
{ Block} & { Statements?} & { Action} & { Evaluated}
\\\hline\hline
\code{user initialization} & n/a & transform & chain
\\[3pt]
\code{random initialization} & n/a & randomize & chain 
\\\hline\hline
\code{data} & no & read & chain  
\\
\code{transformed data} & yes & evaluate & chain  
\\ \hline
\code{parameters} & no & inv.\ transform, Jacobian & leapfrog  \\
& & inv.\ transform, write & sample 
\\[3pt]
\code{transformed parameters} & yes & evaluate & leapfrog \\
& & write & sample 
\\\hline
\code{model} & yes & evaluate & leapfrog
\\\hline
\code{generated quantities} & yes & evaluate & sample \\
& & write & sample
\end{tabular}
\end{center}
\caption{\it Each \proglang{Stan} program block admits certain actions that are evaluated
  at specific times during sampling.  For example, the parameter initialization in
  the \code{data} block requires a read operation once per chain.}
\label{block-actions.fig}
\end{figure}
%
This table is defined assuming HMC or NUTS, both of which apply the
leapfrog integrator to simulate the Hamiltonian dynamics
\citep{HoffmanGelman:2014,Neal:2011} of a particle representing a
parameter.  Each leapfrog step requires an evaluation of the log
density and its gradient.  NUTS automatically adapts the integration
time (and hence number of leapfrog steps) each iteration and uses slice
sampling for improved acceptance, whereas basic HMC sets integration
time statically in the command-line parameters using
\code{engine=static int\_time=t}, with \code{t} set to some positive
floating point value.
%
\begin{figure}
\begin{center}
\begin{tabular}{l|l}
{ Variable Categorization} & { Declaration Block}
\\ \hline\hline
% constants & \code{transformed data}
% \\ \hline
unmodeled data & \code{data}, \code{transformed data}
\\ 
modeled data & \code{data}, \code{transformed data}
\\ \hline
missing data & \code{parameters}, \code{transformed parameters}
\\
modeled parameters & \code{parameters}, \code{transformed parameters}
\\
unmodeled parameters & \code{data}, \code{transformed data}
\\[2pt] \hline
generated quantities & \code{transformed data}, \code{transformed parameters}, 
\\ 
& \code{generated quantities}
\\ \hline\hline
loop indices & any loop statement
\\ \hline
local variables & any statement block
\\ 
\end{tabular}
\end{center}
\caption{\it Variables of each categorization must be declared in
  specific blocks.  Data may also be expressed using numeric
  literals.}\label{variable-kinds.fig}
\end{figure}
%
\citet[p.~366]{GelmanHill:2007} provide a taxonomy of the kinds of
variables used in Bayesian models.  Figure~\ref{variable-kinds.fig}
contains Gelman and Hill's taxonomy aligned with the corresponding
locations of declarations and definitions in \proglang{Stan}.
Unmodeled data variables includes size constants and regression
predictors.  Modeled data variables include known outcomes or
measurements.  A data variable or constant literal is modeled in a
\proglang{Stan} program if it (or a variable that depends on it)
occurs on the left-hand side of a sampling statement.  Unmodeled
parameters are provided as data, and are either known or fixed to some
value for convenience, a typical use being the parameters of a weakly
informative prior for a parameter.

A modeled parameter is given a distribution by the model (usually
dependent on data and correlated with other parameters).  This can be
done either by placing it (or a variable that depends on it) on the
left-hand side of a sampling statement or by declaring the variable
with a constraint (see Section~\ref{implicit-prior.section}).

Any variable that occurs in a (transformed) data block can also be
provided instead as a constant.  So a user may decide to write \code{y
~ normal(0,1)} or to declare \code{mu} and \code{sigma} as data, and
write \code{y ~ normal(mu,sigma)}.  The latter choice allows the
parameters to be changed without recompiling the model, but requires
them to be specified as part of the input data.

Missing data is a new variable type included here.  In order to
perform inference on missing data, it must be declared as a parameter
and modeled; see \citep{GelmanEtAl:2013} for a discussion of
statistical models of missing data.  Unlike \proglang{BUGS}, data may
not contain a mixture of missing (i.e., \code{NA}) values and numeric
values.  \proglang{Stan} uses the built-in \proglang{C++}
floating-point (and integer) arithmetic, for which there is no
equivalent of \code{NA}.  See \citep{Stan:2013} for strategies for
coding missing data problems in \proglang{Stan}.

Generated quantities include simple transforms of data required for
printing.  For example, the variable of interest might be a standard
deviation, resulting from transforming a precision parameter $\tau$ to
a scale parameter $\sigma = \tau^{-1/2}$.  A non-linearly transformed
variable will have different effective sample sizes and $\hat{R}$
statistics than the variable it is derived from, so it is convenient
to define variables of interest in the generated quantities block to
calculate these statistics automatically.  As seen in the example in
Figure~\ref{hier-binom.fig}, generated quantities may also be used for
event probability estimates.

The generated quantities block may also be used for forward
simulations, generating values to make predictions or to perform
posterior predictive checks; see \citep{GelmanEtAl:2013} for more
information.  The generated quantities block is the only location in
\proglang{Stan} in which random-number generators may be applied
explicitly (they are implicit in parameters).

Calculations in the generated quantities block do not impact the
estimates of the parameters.  In this way, generated quantities can be
used for predictions with a behavior not unlike that of cut in
\proglang{BUGS} \citep{LunnEtAl:2012}.  Nevertheless, we recommend
full Bayesian inference in general, and only use the generated
quantities block for efficiency and clarity of code.  For example, in
many models, generating predictions for unseen data or replicating
data for posterior predictive checks does not affect the estimation of
the parameters of interest.

The list in Figure~\ref{variable-kinds.fig} is completed with two
types of local variables, loop indices and traditional local
variables.  Unlike \proglang{BUGS}, \proglang{Stan} allows local
variables to be declared and assigned.  For example, it is possible to
compute the sum of the squares of entries in an array or vector
\code{y} as follows.%
%
\footnote{The built-in squared norm function is the right way to
  implement sums of squares in \proglang{Stan}.}
%
\begin{Code}
{ 
  real sum_of_squares;
  sum_of_squares <- 0;
  for (n in 1:N)
    sum_of_squares <- sum_of_squares + y[n] * y[n];
}  
\end{Code}
%
Local variables may not be declared with constraints, because there is
no location at which it makes sense to test that they satisfy the
constraints.

===================================================

\subsection{The non-centered parameterization for hierarchical models}\label{centering.section}

Consider the following hierarchical logistic regression model fragment.
%
\begin{Code}
data {
  int<lower=1,upper=K> group[N];  // group of data item n
}  
parameters {
  ...
  vector mu;             // mean coeff
  real<lower=0> sigma;   // coeff scale
  vector[K] beta;        // coeff for group k
}
model {
  beta ~ normal(mu,sigma);
  for (n in 1:N)
    y[n] ~ bernoulli_logit(beta[group[n]] * x[n]);
  ...
}
\end{Code}
%
With this parameterization, the \code{beta[k]} are centered around
\code{mu}.  As a result, a change in \code{sigma} is amplified by the
lower-level parameters and introduces a large change in density across
the posterior.  Unfortunately the expected density variation of an HMC
transition is limited%
%
\footnote{The limited density variation is ultimately a consequence of
  a constant metric.  An additional benefit of RMHMC are transitions
  that can cover much larger variations in density, making it uniquely
  suited to these models; see \citep{Neal:2011}.}
%
and many transitions are required to explore the full posterior.  The
resulting sampler devolves into a random walk with high
autocorrelations.  \cite{BetancourtGirolami:2013} provide a detailed
analysis of the benefits of centered vs.\ non-centered
parameterizations of hierarchical models in Euclidean and Riemannian
HMC, with the conclusion that non-centered are best when data are
sparse and centered when the data strongly identifies coefficients.

The following reparameterization employs a non-centered
parameterization for comparison.
%
\begin{Code}
parameters {
  ...
  vector[K] beta_raw;  // replaces beta as parameter  
}
transformed parameters {
  vector[K] beta;      // beta now a transformed parameter
  beta <- mu + sigma * beta_raw;
}
model {
  beta_raw ~ normal(0,1);  // beta now sampled from unit normal
  ...
}
\end{Code}
%
This reparameterization removes the particular correlations amongst the
hierarchical variables that would otherwise limit the effectiveness
of the samplers.  We recommend starting with the more natural centered
parameterization and moving to the non-centered parameterization if
sampling of the coefficients does not mix well.%
%
\footnote{As the saying goes in computer science, it is easier to
  optimize a correct program than debug an optimized program.}

=============================================

\subsubsection{Metropolis-Hastings}

Random-walk Metropolis-Hastings samplers with multivariate normal
proposals have been prototyped, but not yet integrated into the
released versions of \proglang{Stan}.  The primary purpose of
implementing Metropolis-Hastings is to provide baselines against which
to measure more efficient approaches to sampling.  The covariance
matrix used can be adapted during warmup, assuming either equal
diagonal covariance, diagonal covariance, or a full covariance matrix.
Metropolis-Hastings does not require any derivatives of the log
probability function, but the lack of gradient information induces
random walk behavior, leading to slow mixing.

\subsubsection{Ensemble samplers}

Two flavors of ensemble samplers are also in the works: an affine invariant
ensemble sampler \citep{GoodmanWeare:2010} and a differential
evolution sampler \citep{TerBraak:2006}.  Ensemble samplers do not
require derivative information, but typically require at least as many
chains to be run as the number of dimensions in order to have full
rank search of the posterior.  They also require slightly different
management of output due to a single sampler instance producing
multiple correlated chains of draws.

\proglang{Stan} will soon have an implementation of Riemannian
manifold Hamiltonian Monte Carlo (RMHMC)
\citep{GirolamiCalderhead:2011}, using the SoftAbs metric to robustly
incorporate posterior curvature into the Hamiltonian simulation
\cite{Betancourt:2012}.  \proglang{Stan}'s MCMC framework also makes
it straightforward to incorporate other metrics.  This implementation
will also generalize NUTS to Riemannian manifolds
\citep{Betancourt:2013}.  Both support adapting step sizes during
warmup.  RMHMC with SoftAbs uses first, second, and third order
derivatives to adapt to the local curvature of the posterior.